\pkg{Rgbp}: An R Package for Gaussian, Poisson, and Binomial Hierarchical Modeling

### 2013-08-05 jss/article.tex
chagnes

When fitting a model it is always a good idea to question how reliable the estimation procedure is. For example, does our procedure generate interval estimates that have good repeated sampling properties? To answer this question the \code{coverage} function generates pseudo-datasets regarding the estimated $r~(=336.37)$ as a given true value. For reference, we can try any other value of $r$, for example $r=200$, by replacing the code below with `\code{R> pcv <- coverage(p, A.or.r = 200, mean.PriorDist = 0.02, nsim = 10000)}.

@@

\begin{CodeChunk}
\begin{CodeInput}
R> pcv <- coverage(p, nsim = 10000)
\end{CodeInput}
\end{CodeChunk}
\begin{figure}[h] 
\begin{center}
\includegraphics[width = 5.5cm]{hospital2.png}
\caption{Coverage plot via frequency method checking for 31 hospitals}
\label{fig:hospitalcoverage}
\end{center}
\end{figure}

In Figure \ref{fig:hospitalcoverage} we see two estimated coverage probabilities: a simple unbiased estimate (a blue dashed line) and an Rao-Blackwellized unbiased estimate (a red line). Both lines are practically indistinguishable from the horizontal line at 0.95 and the estimated overall average coverage rate is 0.950 from the Rao-Blackwellized estimate. This result shows that the interval estimates for this particular dataset accurately achieves 95\% confidence under repeated sampling.

% For reference, it took 10 seconds for MacBook Pro with a 2.3 GHz dual-core Intel i5 CPU to run 10,000 simulations.

@@

\begin{CodeChunk}
\begin{CodeInput}
R> data(schools)
R> attach(schools)
R> g <- gbp(y, se, model = "gr")
R> g
\end{CodeInput}
\begin{CodeOutput}
Summary for each unit (sorted by se):

@@

\begin{CodeChunk}
\begin{CodeInput}
R> gcv <- coverage(g, nsim = 10000)
\end{CodeInput}
\end{CodeChunk}
\begin{figure}[h] 
\begin{center}
\includegraphics[width = 5.5cm]{school2.png}
\caption{Coverage plot via frequency method checking for 8 schools}
\label{fig:schoolcoverage}
\end{center}
\end{figure}

As seen in Figure \ref{fig:schoolcoverage} the desired $95\%$ confidence (horizontal line at 0.95) is achieved (actually, exceeded) for each school in this example.

@@

Then, how much can we trust these interval estimates? The following frequency method checking that regards the estimated values, $\hat{r}$ (=112.95) and $\hat{\beta}~(=(-1.194, ~0.389)^{T})$, as true values when it generates pseudo-datasets will answer it. For reference, if we want to try different true parameters, the appropriate code will be \code{coverage(b, A.or.r = 100, reg.coef = c(-1, 0.2), nsim = 10000)} instead of below one.

\begin{CodeChunk}
\begin{CodeInput}
R> bcv <- coverage(b, nsim = 10000)
\end{CodeInput}
\end{CodeChunk}
\begin{figure}[h]
\begin{center}
\includegraphics[width = 5.5cm]{baseball2.png}
\caption{Coverage plot via frequency method checking for 18 players}
\label{fig:baseball2}
\end{center}
\end{figure}

Finally, in Figure \ref{fig:baseball2}, we see that the estimated average coverage probability is 0.973 based on the Rao-Blackwellization, conservatively satisfying the definition of the 95\% confidence interval.

% As for the running time, it took 243 seconds for MacBook Pro equipped with a 2.3 GHz dual-core Interl i5 CPU to execute 10,000 simulations.

>>>>

When fitting a model it is always a good idea to question how reliable the estimation procedure is. For example, does our procedure generate interval estimates that have good repeated sampling properties? To answer this question the \code{coverage} function generates pseudo-datasets regarding the estimated $r~(=336.37)$ as a given true value. For reference, we can try any other value of $r$, for example $r=200$, by replacing the code below with `\code{R> pcv <- coverage(p, A.or.r = 200, mean.PriorDist = 0.02, nsim = 1000)}.

@@

\begin{CodeChunk}
\begin{CodeInput}
R> pcv <- coverage(p, nsim = 1000)
\end{CodeInput}
\end{CodeChunk}
\begin{figure}[h] 
\begin{center}
\includegraphics[width = 5.5cm]{hospital2.png}
\caption{Coverage plot via frequency method checking for 31 hospitals}
\label{fig:hospitalcoverage}
\end{center}
\end{figure}

In Figure \ref{fig:hospitalcoverage} we see three lines; one is a red line connecting Rao-Blackwellized unbiased estimates for 31 hospitals; another is a dashed blue line connecting simple unbiased estimates for those hospitals; a black horizontal line indicating the confidence level, 0.95.

The estimated overall average coverage rate is 0.950 from both Rao-Blackwellized and simple unbiased estimates. This result shows that the interval estimates for this particular dataset accurately achieves 95\% confidence under repeated sampling. Note that this estimate depends on the given true value of $r$ and the assumption that the model is true.

If we want to check average Rao-Blackwellized unbiased estimates of coverage rate for each hospital, the following code gives us 31 estimates.
\begin{CodeChunk}
\begin{CodeInput}
R> pcv$coverageRB
\end{CodeInput}
\begin{CodeOutput}
 [1] 0.950 0.950 0.950 0.951 0.950 0.950 0.951 0.950 0.950 0.950 0.951 0.951 
[13] 0.950 0.950 0.950 0.950 0.951 0.950 0.950 0.950 0.950 0.951 0.950 0.950 
[25] 0.951 0.950 0.951 0.950 0.951 0.951 0.950
\end{CodeOutput}
\end{CodeChunk}

As for average simple unbiased estimates of coverage rate for each hospital, we can type the code below to obtain 31 estimates. 
\begin{CodeChunk}
\begin{CodeInput}
R> pcv$coverageS
\end{CodeInput}
\begin{CodeOutput}
 [1] 0.949 0.953 0.946 0.955 0.957 0.939 0.956 0.957 0.944 0.952 0.956 0.955 
[13] 0.949 0.957 0.948 0.954 0.951 0.934 0.956 0.943 0.959 0.952 0.949 0.964 
[25] 0.949 0.953 0.941 0.938 0.941 0.956 0.951
\end{CodeOutput}
\end{CodeChunk}

The last thing to mention is the accuracy of Rao-Blackwellized unbiased estimates compared to that of simple one. Taking the first hospital as an example, we can calculate the sample variance of 1,000 Rao-Blackwellized unbiased estimates for the first hospital and that of 1,000 simple ones by typing following code. For reference, \code{raw.resultRB} and \code{raw.resultS} in \code{pcv} contains all the individual simulation results.
\begin{CodeChunk}
\begin{CodeInput}
R> var(pcv$raw.resultRB[1, ])
\end{CodeInput}
\begin{CodeOutput}
 [1] 0.002031225
\end{CodeOutput}
\begin{CodeInput}
R> var(pcv$raw.resultS[1, ])
\end{CodeInput}
\begin{CodeOutput}
 [1] 0.04844745
\end{CodeOutput}
\end{CodeChunk}

It shows that Rao-Blackwellized unbiased estimate is about 24 times more accurate than simple one. Also, as the number of simulation increases (for example 10,000), the zig-zagging blue line will be closer to the red line.
% For reference, it took 10 seconds for MacBook Pro with a 2.3 GHz dual-core Intel i5 CPU to run 10,000 simulations.

@@

\begin{CodeChunk}
\begin{CodeInput}
R> data(schools)
R> g <- gbp(y, se, model = "gr")
R> g
\end{CodeInput}
\begin{CodeOutput}
Summary for each unit (sorted by se):

@@

\begin{CodeChunk}
\begin{CodeInput}
R> gcv <- coverage(g, nsim = 1000)
\end{CodeInput}
\end{CodeChunk}
\begin{figure}[h] 
\begin{center}
\includegraphics[width = 5.5cm]{school2.png}
\caption{Coverage plot via frequency method checking for 8 schools}
\label{fig:schoolcoverage}
\end{center}
\end{figure}

As seen in Figure \ref{fig:schoolcoverage} the desired $95\%$ confidence (horizontal line at 0.95) is achieved (actually, exceeded) for each school in this example. Here, all the coverage estimates depend on the given true values of $A$ and $\beta_{0}$, and the assumption that the model is true.

Average Rao-Blackwellized unbiased estimates of coverage rate for each school can be seen as below.
\begin{CodeChunk}
\begin{CodeInput}
R> gcv$coverageRB
\end{CodeInput}
\begin{CodeOutput}
 [1] 0.966 0.957 0.966 0.962 0.960 0.961 0.960 0.969
\end{CodeOutput}
\end{CodeChunk}

As for average simple unbiased estimates of coverage rate for each school, 
\begin{CodeChunk}
\begin{CodeInput}
R> gcv$coverageS
\end{CodeInput}
\begin{CodeOutput}
[1] 0.970 0.943 0.957 0.968 0.956 0.969 0.961 0.976
\end{CodeOutput}
\end{CodeChunk}

Finally  when it comes to the first school,  Rao-Blackwellized unbiased estimate is about 24 times more accurate than simple one. It is based on the ratio of the sample variances of individual 1,000 estimates from two approaches.
\begin{CodeChunk}
\begin{CodeInput}
R> var(pcv$raw.resultS[1, ]) / var(pcv$raw.resultRB[1, ])
\end{CodeInput}
\begin{CodeOutput}
 [1] 23.85134
\end{CodeOutput}
\end{CodeChunk}

@@

Then, how much can we trust these interval estimates? The following frequency method checking that regards the estimated values, $\hat{r}$ (=112.95) and $\hat{\beta}~(=(-1.194, ~0.389)^{T})$, as true values when it generates pseudo-datasets will answer it. For reference, if we want to try different true parameters, the appropriate code will be \code{coverage(b, A.or.r = 100, reg.coef = c(-1, 0.2), nsim = 1000)} instead of below one.

\begin{CodeChunk}
\begin{CodeInput}
R> bcv <- coverage(b, nsim = 1000)
\end{CodeInput}
\end{CodeChunk}
\begin{figure}[h]
\begin{center}
\includegraphics[width = 5.5cm]{baseball2.png}
\caption{Coverage plot via frequency method checking for 18 players}
\label{fig:baseball2}
\end{center}
\end{figure}

Finally, in Figure \ref{fig:baseball2}, we see that the estimated average coverage probability is 0.972 based on the Rao-Blackwellization, conservatively satisfying the definition of the 95\% confidence interval. Each coverage estimate depends on given true values of $r$ and $\beta_{(2\times1)}$, and the assumption that the model is true.

The average Rao-Blackwellized unbiased estimates of coverage rate for each player follows.
\begin{CodeChunk}
\begin{CodeInput}
R> bcv$coverageRB
\end{CodeInput}
\begin{CodeOutput}
 [1] 0.972 0.972 0.973 0.972 0.972 0.970 0.974 0.972 0.973 0.973 0.973 0.970 
[13] 0.972 0.972 0.973 0.972 0.973 0.972
\end{CodeOutput}
\end{CodeChunk}

And also average simple unbiased estimates of coverage rate for each player can be obtained.
\begin{CodeChunk}
\begin{CodeInput}
R> bcv$coverageS
\end{CodeInput}
\begin{CodeOutput}
 [1] 0.969 0.967 0.973 0.980 0.970 0.974 0.965 0.976 0.969 0.972 0.980 0.972 
[13] 0.957 0.969 0.972 0.982 0.978 0.984
\end{CodeOutput}
\end{CodeChunk}

Finally,  we see that Rao-Blackwellized unbiased estimate for the first player is about 26 times more accurate than simple one. It is based on the ratio of the sample variances of individual estimates from two approaches.
\begin{CodeChunk}
\begin{CodeInput}
R> var(bcv$raw.resultS[1, ]) / var(bcv$raw.resultRB[1, ])
\end{CodeInput}
\begin{CodeOutput}
 [1] 25.64346
\end{CodeOutput}
\end{CodeChunk}
% As for the running time, it took 243 seconds for MacBook Pro equipped with a 2.3 GHz dual-core Interl i5 CPU to execute 10,000 simulations.
